---
title: "UQ R Markdown template"
author: "[Yan Holtz](https://github.com/holtzy)"
date: "`r format(Sys.time(), '%d %B %Y')`"
mail: "yan.holtz.data@gmail.com"
linkedin: "yan-holtz-2477534a"
twitter: "r_graph_gallery"
github: "holtzy"
home: "www.yan-holtz.com"
output:
  epuRate::UQ:
    toc: TRUE
    number_sections: FALSE
    code_folding: "hide"
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# You need these libraries to run this template:
library(rmarkdown)    # install.packages("rmarkdown")
library(epuRate)      # devtools::install_github("holtzy/epuRate", force=TRUE)
```

<br><br><br><br>

>Welcome in the [PCTG R Markdown template](https://github.com/holtzy/epuRate). This document describes how the template looks like, building an interactive Manhattan plot.






# Loading data
***
Let's load a GWAS summary statistic stored in the [qqman library](https://github.com/stephenturner/qqman). This library is:

- Made by Stephen Turner
- Good to plot a quick Manhattan plot

```{r, message=FALSE}
library(qqman)
```






# Show data in a table
***
It's often handy to keep a trace of the raw data somewhere in your document. The DT library allows to build interactive tables that you can search, filter, highlight and more.
```{r}
library(DT)
datatable(gwasResults, rownames = FALSE, filter="top", options = list(pageLength = 5, scrollX=T) )
```





# Manhattan plot
***
Using `HTML` outputs you can embed some interactive graphics. For example, the plotly library can transform any of your ggplot2 graphic in an interactive chart:
```{r, message=FALSE, warning=FALSE, echo=FALSE}
# Libraries
library(plotly)
library(tidyverse)

# Prepare the dataset
don <- gwasResults %>%

  # Compute chromosome size
  group_by(CHR) %>%
  summarise(chr_len=max(BP)) %>%

  # Calculate cumulative position of each chromosome
  mutate(tot=cumsum(chr_len)-chr_len) %>%
  select(-chr_len) %>%

  # Add this info to the initial dataset
  left_join(gwasResults, ., by=c("CHR"="CHR")) %>%

  # Add a cumulative position of each SNP
  arrange(CHR, BP) %>%
  mutate( BPcum=BP+tot) %>%

  # Add highlight and annotation information
  mutate( is_highlight=ifelse(SNP %in% snpsOfInterest, "yes", "no")) %>%

  # Filter SNP to make the plot lighter
  filter(-log10(P)>0.5)

# Prepare X axis
axisdf <- don %>% group_by(CHR) %>% summarize(center=( max(BPcum) + min(BPcum) ) / 2 )

# Prepare text description for each SNP:
don$text <- paste("SNP: ", don$SNP, "\nPosition: ", don$BP, "\nChromosome: ", don$CHR, "\nLOD score:", -log10(don$P) %>% round(2), "\nWhat else do you wanna know", sep="")
```

```{r, message=FALSE, warning=FALSE, fig.width=9}
# Make the plot
p <- ggplot(don, aes(x=BPcum, y=-log10(P), text=text)) +

    # Show all points
    geom_point( aes(color=as.factor(CHR)), alpha=0.8, size=1.3) +
    scale_color_manual(values = rep(c("grey", "#563986"), 22 )) +

    # custom X axis:
    scale_x_continuous( label = axisdf$CHR, breaks= axisdf$center ) +
    scale_y_continuous(expand = c(0, 0) ) +     # remove space between plot area and x axis

    # Add highlighted points
    geom_point(data=subset(don, is_highlight=="yes"), color="orange", size=2) +

    # Custom the theme:
    theme_bw() +
    theme(
      legend.position="none",
      panel.border = element_blank(),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank()
    ) +
    ylim(0, 10)

ggplotly(p, tooltip="text")
```
